\LARGE A double-sampling extension of the German National Forest Inventory for design-based small area regression estimation of timber volume resources on forest district levels \\ \vspace{2ex} \large Large scale application to the federal state of Rhineland-Palatinate \vspace{-1ex}

### 2018-02-05 Draft_Form/tex/Results.tex
update 05_02_2018
update 05_02_2018

An application of the SRS, \psmall{} and \extpsynth{} estimator was not feasible for 17 of all 405 FR-units due to an insufficient terrestrial sample size of $n_{2,G} < 2$. We further restricted the calculation of \psmall{} and \extpsynth{} estimator to small area units with a minimum terrestrial sample size of $n_{2,G} \geq 4$ to avoid unstable estimates. This affected 65 additional FR units and limited unbiased two-phase estimations to 321 (79\%) of the 405 FR units. It should be noted that also the \psynth{} estimator could not be applied for 2 FR-units since $n_{1,G} < 2$. Due to substantial larger sample sizes, all estimators could however be applied in all small area units on the FA level. The average value and the range of the mean timber volume estimates over the evaluated FA and FR units turned out to be very similar between all estimators (table \ref{tab:estres}). An additional pairwise comparison of the 95\% confidence intervals revealed that the four estimators did in fact not produce statistically different point estimates for all FA and FR units. This confirmed that the differences between the estimators are solely found in the precision which they provide for the point estimates.

@@

On both small area levels, the \psynth{} estimator resulted in much smaller estimation errors compared to \psmall{} and \extpsynth{}. This was as expected, since the \psynth{} variance estimate does not take the residual variation in each small area unit into account (section \ref{sec:psynth}). Compared to the asymptotically design-unbiased estimators \psmall{} and \extpsynth{}, the estimation errors produced by \psynth{} seem to be too optimistic. One should also recall that the estimates of the \psynth{} estimator are potentially design-biased.

@@

Figure \ref{fig:disterrors} reveals that the error distribution of \psmall{} and \extpsynth{} are very similar, with \psmall{} showing marginally higher estimation errors. In order to investigate the differences between \psmall{} and \extpsynth{}, we compared the g-weight variances of both estimators for all 321 FR units (fig. \ref{fig:compvar}, left). As obvious, \psmall{} yielded slightly larger variances for the vast majority of the estimates. As addressed in section \ref{sec:extpsynth}, one possible explanation for such differences was the effect of one or more cluster not entirely being included in a small area unit, as this would constitute a violation of the \extpsynth{} estimator. This violation was actually observed in 155 of the 321 FR units (48\%). However, the affected FR units (depicted in red diamonds, fig. \ref{fig:compvar}) did not show a significant divergence from the \psmall{} variances with respect to the remaining unaffected FR units. The variance differences between the two estimators were thus due to the mathematical formulations of the \psmall{} and \extpsynth{} estimator, which are asymptotically equivalent only under large terrestrial sample sizes $n_{2,G}$ within the small area \citep[pp.17--18]{mandallaz2016}. An additional comparison of the absolute differences in the g-weight variance (fig. \ref{fig:compvar}, right) revealed that large divergences did in fact particularly occur for small area units with small terrestrial sample sizes ($n_{2,G} \leq 5$). The differences decreased with increasing sample size and thus confirmed the asymptotic relationship between the two estimators. However, a comparison of the confidence intervals of \psmall{} and \extpsynth{} revealed that the variance differences did not lead to statistically significant point estimates. The results indicate that the adjustment of the intercept term of the \extpsynth{} estimator to the data within a small area will in general yield slightly smaller variances than the \psmall{} estimator.\par

@@

% ----------------------------------------------------------------------- %
% ----------------------------------------------------------------------- %
\subsection{Reduction of SRS variance by double sampling procedure}
\label{sec:gain_eval}

A direct comparison of the achieved variances within the small area units revealed that the application of the design-unbiased estimators (\psmall{} and \extpsynth{}) led to a reduction of the SRS variance in all FA units. In 75\% of the FA units, the \extpsynth{} estimator was able to reduce the SRS variance by up to 54.1\% (fig. \ref{fig:gain}). The reduction in variance was also expressed in the relative efficiency values, which were 2.02 on average and ranged between 1.18 and 4.13 on the FA level. On FR-level, the achieved reduction in variance and the relative efficiencies reached even higher values (table \ref{tab:gain}). The \psmall{} estimator again yielded slightly lower variance reductions and relative efficiencies due to the smaller variances produced by the \extpsynth{} estimator (section \ref{sec:comp}). However, cases also occurred where one or both two-phase estimators produced substantially larger variance values than under the SRS estimator. This particularly happened in 19\% (61) of the FR units under the \extpsynth{}, and in 24\% (76) of the FR units under the \psmall{} estimator.

@@

One possible reason for an increased variance by the two-phase estimators was supposed to be a poor performance of the regression model within the small area unit. In order to investigate this hypothesis, we analyzed the three variance terms of the  \psmall{} estimator (eq. \ref{eq:psmall_var}), i.e. the variance introduced by the uncertainty of the regression coefficients (term 1), the variance caused by estimating the auxiliary means (term 2), and the variance of the model residuals that expresses the model performance within the small area (term 3). Figure \ref{fig:fail1} illustrates the reduction or increase of the SRS variance for all FR units in dependence of the proportion of the residual variance term (term 3) on the g-weight variance. Two issues become obvious: firstly, the residual term generally constitutes the dominating part of the g-weight variance (around 84\% on average). FR units for which the \psmall{} estimator produced an increase of the SRS variance did not differ distinctly from the pattern of the remaining FR units. The same was also found when looking at the absolute residual terms (fig. \ref{fig:fail2}). Secondly, FR units with the highest increases of the SRS variance were particularly observed under small terrestrial sample sizes $n_{2,G} = 4$. The average increase in variance for $n_{2,G} = 4$ was 272\% compared to 62\% for $n_{2,G} > 4$. In comparison, the average reduction in SRS variances was around 50\% for both sampling sizes $n_{2,G} = 4$ and $n_{2,G} > 4$. These findings suggest that moderate or poor model fits within a small area can cause the \psmall{} and \extpsynth{} estimator to produce higher variances than the SRS estimator particularly under small terrestrial sample sizes. However, small terrestrial sample sizes can also have the effect that the SRS estimator might not reflect the actual variation of the local density within a small area. In this case, the two-phase variance estimate might be larger but more realistic. Whereas a visual analysis of aerial images, remote sensing data or stand maps might give some evidence for this hypothesis, a definite proof is practically infeasible.

>>>>

An application of the SRS, \psmall{} and \extpsynth{} estimator was not feasible for 17 of all 405 FR-units due to an insufficient terrestrial sample size of $n_{2,G} < 2$. We further restricted the calculation of the \psmall{} and \extpsynth{} estimator to small area units with a minimum terrestrial sample size of $n_{2,G} \geq 4$ to avoid unstable estimates. This affected 65 additional FR units and limited unbiased two-phase estimations to 321 (79\%) of the 405 FR units. It should be noted that also the \psynth{} estimator could not be applied for 2 FR-units since $n_{1,G} < 2$. Due to substantial larger sample sizes, all estimators could however be applied to all 45 FA units. The average value and the range of the mean timber volume estimates over the evaluated FA and FR units turned out to be very similar between all estimators (table \ref{tab:estres}). An additional pairwise comparison of the 95\% confidence intervals revealed that the four estimators did in fact not produce statistically different point estimates for all FA and FR units. This confirmed that the differences between the estimators are solely found in the precision which they provide for the point estimates.

@@

On both small area levels, the \psynth{} estimator resulted in much smaller estimation errors compared to \psmall{} and \extpsynth{}. This was as expected, since the \psynth{} variance estimate does not take the residual variation in each small area unit into account (section \ref{sec:psynth}). Compared to the asymptotically design-unbiased estimators \psmall{} and \extpsynth{}, the estimation errors produced by \psynth{} thus seem to be too optimistic. One should also recall that the estimates of the \psynth{} estimator are potentially design-biased.

@@

Figure \ref{fig:disterrors} reveals that the error distribution of \psmall{} and \extpsynth{} are very similar, with \psmall{} showing marginally higher estimation errors. In order to investigate the differences between \psmall{} and \extpsynth{}, we compared the g-weight variances of both estimators for all 321 FR units (fig. \ref{fig:compvar}, left). As obvious, \psmall{} yielded slightly larger variances for the vast majority of the estimates. As addressed in section \ref{sec:extpsynth}, one possible explanation for such differences was the effect of one or more cluster not entirely being included in a small area unit, as this would constitute a violation of the \extpsynth{} estimator. This violation was actually observed in 155 of the 321 FR units (48\%). However, the affected FR units (depicted in red diamonds, fig. \ref{fig:compvar}) did not show a significant divergence from the \psmall{} variances with respect to the remaining unaffected FR units. The variance differences between the two estimators were thus due to the mathematical formulations of the \psmall{} and \extpsynth{} estimator, which are asymptotically equivalent only under large terrestrial sample sizes $n_{2,G}$ within the small area \citep[pp.17--18]{mandallaz2016}. An additional comparison of the absolute differences in the g-weight variance (fig. \ref{fig:compvar}, right) revealed that large divergences did in fact particularly occur for small area units with small terrestrial sample sizes ($n_{2,G} \leq 5$). The differences decreased with increasing sample size and thus confirmed the asymptotic relationship between the two estimators. However, a comparison of the confidence intervals of \psmall{} and \extpsynth{} revealed that the variance differences did not lead to statistically significant point estimates. The results indicate that the adjustment of the intercept term to the data within a small area used in the \extpsynth{} estimator will in general yield slightly smaller variances than the \psmall{} estimator.\par

@@

% ----------------------------------------------------------------------- %
% ----------------------------------------------------------------------- %
\subsection{Reduction of SRS variance by \psmall{} and \extpsynth{}}
\label{sec:gain_eval}

A direct comparison of the achieved variances within the small area units revealed that the application of the design-unbiased estimators (\psmall{} and \extpsynth{}) led to a reduction of the SRS variance in all FA units. In 75\% of the FA units, the \extpsynth{} estimator was able to reduce the SRS variance by up to 54.1\% (fig. \ref{fig:gain}). The reduction in variance was also expressed in the relative efficiency values, which were 2.02 on average and ranged between 1.18 and 4.13 on the FA level. On FR-level, the reduction in variance and the relative efficiencies reached even higher values (table \ref{tab:gain}). The \psmall{} estimator again yielded slightly lower variance reductions and relative efficiencies due to the smaller variances produced by the \extpsynth{} estimator (section \ref{sec:comp}). However, cases also occurred on the FR level where one or both two-phase estimators produced larger variance values than under the SRS estimator. This particularly happened in 19\% (61) of the FR units under the \extpsynth{}, and in 24\% (76) of the FR units under the \psmall{} estimator.

@@

One possible reason for an increased variance by the two-phase estimators was supposed to be a poor performance of the regression model within the small area unit. In order to investigate this hypothesis, we analyzed the three variance terms of the  \psmall{} estimator (eq. \ref{eq:psmall_var}), i.e. the variance introduced by the uncertainty of the regression coefficients (term 1), the variance caused by estimating the auxiliary means (term 2), and the variance of the model residuals that expresses the model performance within the small area (term 3). Figure \ref{fig:fail1} illustrates the reduction or increase of the SRS variance by \psmall{} for all FR units in dependence of the residual variance term proportion on the g-weight variance. Two issues become obvious: firstly, the residual term generally constitutes the dominating part of the g-weight variance for all \psmall{} estimates (around 84\% on average). FR units for which the \psmall{} estimator produced larger variances than the SRS estimator did not differ distinctly from the pattern of the remaining FR units. The same was also found when looking at the absolute residual term values (fig. \ref{fig:fail2}). Secondly, FR units with the highest variance increases compared to SRS were particularly observed under small terrestrial sample sizes $n_{2,G} = 4$. The average increase in variance for $n_{2,G} = 4$ was 272\% compared to 62\% for $n_{2,G} > 4$. In comparison, the average reduction in SRS variances was around 50\% for both sampling sizes $n_{2,G} = 4$ and $n_{2,G} > 4$. These findings suggest that moderate or poor model fits within a small area can cause the \psmall{} and \extpsynth{} estimator to produce higher variances than the SRS estimator particularly under small terrestrial sample sizes. However, small terrestrial sample sizes can also have the effect that the SRS estimator might not reflect the actual variation of the local density within a small area. In this case, the two-phase variance estimate might be larger but more realistic. Whereas a visual analysis of aerial images, remote sensing data or stand maps might give some evidence for this hypothesis, a definite proof is practically infeasible.
